Solving Collective Spin Hamiltonians with Decoherence

### 2018-10-02 soc_squeezing/cumulant_expansions.tex
more simplifying of the operator product, and code to test it

We wish to change our basis for all exchange-symmetric operators on a
system of $N$ qubits from $S_+^\ell S_\z^m S_-^n$ to
\begin{align}
  \S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
  &&
  P_{\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j_a^\mu}},
\end{align}
where $\v m \equiv \p{m_+,m_\z,m_-}$; $\v j$ is an ordered list of
indices for the spins addressed by $P_{\v j}$; and the sum over
$\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for which all
indices in $\v j$ are distinct and $\abs{\set{j_a^\mu}} = m_\mu$.  In
words, $\S_{\v m}$ is an exchange-symmetric $\p{m_++m_\z+m_-}$-spin
operator with $m_\mu$ single-spin operators of type $\sigma_\mu$, and
$P_{\v j}$ is a single term in $\S_{\v m}$ which addresses the spins
indexed by $\v j$.

We now expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$
into terms that have exactly $s$ spins addressed by both operators:
\begin{align}
  \S_{\v m} \S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
  &&
  \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
  \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X,
  \label{eq:SS_PP}
\end{align}
where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
in $P_{\v j}$ address the same spin as a $\sigma_\nu$ operator in
$P_{\v k}$, we have
\begin{align}
  \S_{\v m}\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v r,\v J,\v K;\v m,\v n,s}}
  g_{\v m\v n\v r} Q_{\v J} \tilde Q_{\v K},
  \label{eq:SS_PQ}
\end{align}
where $\v J$ and $\v K$ are respectively ordered lists of indices for
spins addressed by one and two single-spin operators in each term of
the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
over $\p{\v r,\v J,\v K;\v m,\v n,s}$ implies a sum over all possible
values of $\v r,\v J,\v K$ with the restrictions
\begin{align}
  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
  &&
  \sum_\nu r_{\mu\nu} \le m_\mu,
  &&
  \sum_\mu r_{\mu\nu} \le n_\nu,
  &&
  \sum_{\mu,\nu} r_{\mu\nu} = s,
\end{align}
\begin{align}
  \abs{\set{J_a^\mu}}
  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
  &&
  \abs{\set{K_a^{\mu\nu}}} = r_{\mu\nu},
  &&
  \set{J_a^\mu} \cap \set{K_a^{\mu\nu}} = \set{};
\end{align}
the factor $g_{\v m\v n\v r}$ counts the number of ways to pair
operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
\begin{align}
  g_{\v m\v n\v r}
  &\equiv \sp{\prod_{\mu,\nu}
    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
\end{align}
where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
and finally
\begin{align}
  Q_{\v J} \equiv \prod_{\mu,a} \sigma_\mu^{(J_a^\mu)},
  &&
  \tilde Q_{\v K} \equiv \prod_{\mu,\nu,a}
  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}
  = \prod_{\mu,\nu,a} \sum_\rho
  \eta_{\mu\nu\rho} \sigma_\rho^{(K_a^{\mu\nu})},
  \label{eq:Q_JK}
\end{align}
where $\eta_{\mu\nu\rho}$ is defined by the product
$\sigma_\mu\sigma_\nu\equiv\sum_\rho\eta_{\mu\nu\rho}\sigma_\rho$.
Note that unlike the sums over $\mu$ and $\nu$ in the work above, the
sum over $\rho$ in \eqref{eq:Q_JK} now includes an index for the
identity operator $\sigma_0\equiv\1$.

>>>>

We wish to change our basis for all exchange-symmetric operators on a
system of $N$ qubits from $S_+^\ell S_\z^m S_-^n$ to
\begin{align}
  \S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
  &&
  P_{\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j_a^\mu}},
\end{align}
where $\v m \equiv \p{m_\z,m_+,m_-}$; $\v j$ is an ordered list of
indices for the spins addressed by $P_{\v j}$; and the sum over
$\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for which all
indices in $\v j$ are distinct and $\abs{\set{j_a^\mu}} = m_\mu$.  In
words, $\S_{\v m}$ is an exchange-symmetric $\p{m_\z+m_++m_-}$-spin
operator with $m_\mu$ single-spin operators of type $\sigma_\mu$, and
$P_{\v j}$ is a single term in $\S_{\v m}$ which addresses the spins
indexed by $\v j$.

We now expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$
into terms that have exactly $s$ spins addressed by both operators:
  \S_{\v m} \S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
  \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
  \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X,
  \label{eq:SS_PP}
where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
in $P_{\v j}$ address the same spin as a $\sigma_\nu$ operator in
$P_{\v k}$, we have
  \S_{\v m}\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v J,\v K,\v r;\v m,\v n,s}}
  f_{\v m\v n\v r} P_{\v J} Q_{\v K},
  \label{eq:SS_PQ}
where $\v J$ and $\v K$ are respectively ordered lists of indices for
spins addressed by one and two single-spin operators in each term of
the expansion for $\S_{\v m} \S_{\v n}$ in \eqref{eq:SS_PP}; the sum
over $\p{\v J,\v K,\v r;\v m,\v n,s}$ denotes a sum over all values of
$\v J,\v K,\v r$ with the restrictions
  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
  \sum_\nu r_{\mu\nu} \le m_\mu,
  \sum_\mu r_{\mu\nu} \le n_\nu,
  \sum_{\mu,\nu} r_{\mu\nu} = s,
  \abs{\set{J_a^\mu}}
  = m_\mu + n_\mu - \sum_\kappa\p{r_{\mu\kappa}+r_{\kappa\mu}},
  \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \set{};
the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
operators in $P_{\v j}$ and $P_{\v k}$ as prescribed by $\v r$, or
  f_{\v m\v n\v r}
  &\equiv \sp{\prod_{\mu,\nu}
    { m_\mu - \sum_{\kappa<\nu} r_{\mu\kappa} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\kappa<\nu} r_{\kappa\mu} \choose r_{\nu\mu} }}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\kappa r_{\mu\kappa}}
    \p{n_\mu}_{\sum_\kappa r_{\kappa\mu}}}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1},
where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
and finally, similarly to the definition of $P_{\v j}$ in
\eqref{eq:SS_PP} we have
  Q_{\v K} \equiv \prod_{\mu,\nu,a}
  \sigma_\mu^{(K^{\mu\nu}_a)} \sigma_\nu^{(K^{\mu\nu}_a)}.
Expanding the product
$\sigma_\mu\sigma_\nu=\sum_\kappa\eta_{\mu\nu\rho}\sigma_\rho$, where
unlike the sums over $\mu,\nu$ in most of the work above, the sum over
$\rho$ here includes an index for the identity operator
$\sigma_0\equiv\1$, we can expand and collect terms to get
  Q_{\v K}
  = \prod_{\mu,\nu,a} \sum_\rho
  \eta_{\mu\nu\rho} \sigma_\rho^{(K^{\mu\nu}_a)}
  = \sum_{\v\rho} \prod_{\mu,\nu,a} \eta_{\mu\nu\rho^{\mu\nu}_a}
  \sigma_{\rho^{\mu\nu}_a}^{(K^{\mu\nu}_a)},
where we have distributed a product of sums into a sum of products.
Collecting together operators of type $\sigma_\kappa$, we then have
  Q_{\v K}
  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,\kappa}
    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\v\rho\kappa}_a)},
where $\tilde\rho_\kappa^{\mu\nu}$ is the number of elements in
$\v\rho^{\mu\nu} \equiv
\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
equal to $\kappa$, and $K^{\v\rho\kappa}_a$ indexes the $a$-th
operator of type $\sigma_\kappa$ for each choice of $\v\rho$, such
that for any fixed $\v\rho$ we have
$\set{K^{\v\rho\kappa}_a}=\set{K^{\mu\nu}_b}$.  Exploiting the
exchange symmetry of $Q_{\v K}$ under the sum in \eqref{eq:SS_PQ}, we
can group together all terms which are equal up to a permutation of
indices in $\v\rho^{\mu\nu}$, i.e.~terms with equal values of
$\tilde\rho_\kappa^{\mu\nu}$, to write
  Q_{\v K}
  \simeq \sum_{\tilde{\v\rho}}
  g_{\tilde{\v\rho}} \sp{\prod_{\mu,\nu,\kappa}
    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
  \prod_{\kappa,a} \sigma_{\kappa}^{(K^{\tilde{\v\rho}\kappa}_a)}
  = \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}}
  \sp{\prod_{\mu,\nu,\kappa}
    \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
  P_{\v K_{\tilde{\v\rho}}},
where $\simeq$ denotes equality up to a permutation of spin indices;
the degeneracy factor for each term is
  g_{\tilde{\v\rho}}
  \equiv \prod_{\mu,\nu,\kappa}
  { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
    \choose \tilde\rho^{\mu\nu}_\kappa }
  = \prod_{\mu,\nu}
  \f{r_{\mu\nu}!}{\prod_\kappa\tilde\rho^{\mu\nu}_\kappa!};
as before $K^{\tilde{\v\rho}\kappa}_a$ indexes the $a$-th operator of
type $\sigma_\kappa$ for each partitioning $\tilde{\v\rho}$, such that
for any fixed $\tilde{\v\rho}$ we have
$\set{K^{\tilde{\v\rho}\kappa}_a}=\set{K^{\mu\nu}_b}$; and
$\v K_{\tilde{\v\rho}}$ is an ordered list of the indices in
$\set{K^{\tilde{\v\rho}\kappa}_a}$.

% \begin{align}
%   \S_{\v m}\S_{\v n}
%   = \sum_{s\ge0} \sum_{\p{\v J,\v K,\v r;\v m,\v n,s}}
%   \sum_{\tilde{\v\rho}}
%   f_{\v m\v n\v r} g_{\tilde{\v\rho}}
%   \sp{\prod_{\mu,\nu,\kappa}
%     \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa}}
%   P_{\v J} P_{\v K_{\tilde{\v\rho}}}
% \end{align}
